\bf\mftitle

### 2011-01-22 doc/jInfer/twostep/twostep.tex
3159165 Create documentation fof TwoStep simplifier - removing unneeded pdfs
formatting paragraphs and verbatims in architecture
3159165 Create documentation fof TwoStep simplifier - more text on twostep module, subfigures

\newcommand{\myscale}{0.74}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\jmodule}[1]{\texttt{\textsl{#1}}}

@@

\section{Using of factory pattern}
\begin{figure}
	\centering\includegraphics[scale=\myscale]{factory_pattern1}
	\caption{Same instance of class returned by NetBeans in two successive inference runs} \label{factory_pattern1}
\end{figure}
\jmodule{TwoStepSimplifier} is divided into three submodules, which are then divided into submodules and so on.
Each submodule is simply class, properly anotated.
User selects proper classes - submodules to work in chain on inference.
Classes are then, in runtime looked up by using NetBeans lookup mechanism (see \cite[p. 16]{arch}).
But NetBeans holds one instance of each class, that may be looked upon.
When user runs inference first time, classes are looked up and used.
Maybe class members are initialised, instances are in some state after inference completes.
Then, user clicks run button again, and same instances of classes are returned by lookups.
Oops, theese are not freshly created instances, using them as they are may cause harm.
The problem is ilustrated on fig. \ref{factory_pattern1}.
\begin{figure}
	\centering\includegraphics[scale=\myscale]{factory_pattern2}
	\caption{Factory solution of singleton lookup class problem} \label{factory_pattern2}
\end{figure}

@@

\section{TwoStep module}
TwoStep simplifier is inspired by \cite{1802522} design. Inference proceeds in two steps:
\begin{enumerate}
	\item clustering of element instances into clusters of (probably) same elements
	\item infering regular expression for each element from examples of element contents taken from all elements in cluster
\end{enumerate}
Main routine of TwoStep simplifer does basically this:
\begin{verbatim}
// 1. cluster elements
final Clusterer<AbstractStructuralNode> clusterer= clustererFactory.create();
clusterer.addAll(initialGrammar);
clusterer.cluster();

@@

// 3. process rules
final ClusterProcessor<AbstractStructuralNode> processor= clusterProcessorFactory.create();
for (Cluster<AbstractStructuralNode> cluster : clusterer.getClusters()) {
  final AbstractStructuralNode node =  processor.processCluster(clusterer, cluster.getMembers() - without sentinels);
  // 4. add to rules
  finalGrammar.add(new Element(node.getContext(),
                   node.getName(),
                   node.getMetadata(),
                   ((Element) node).getSubnodes(),
                   attList);
}
\end{verbatim}
Task of clustering is dedicated to \jmodule{Clusterer} submodule, and task of infering regular expression for each cluster is dedicated to \jmodule{ClusterProcessor} submodule. We will examine both of them.
There is also third submodule called \jmodule{RegularExpressionCleaner} actually, its purpose is just to beatify output regular expressions, no inference logic is implemented there. 
Modules currently in jInfer are drawn on fig. \ref{twostep_modules}.
\begin{figure}
	\centering\includegraphics[scale=\myscale]{twostep_modules}
	\caption{Submodules of TwoStep simplifier} \label{twostep_modules}
\end{figure}
We provide one \jmodule{Clusterer} and four \jmodule{ClusterProcessor} implementations.
Each of those will be explained further in this document.

TODO sentinel processing

\subsection{Clusterer submodule}
Lets start by example, let input document(s) contain XML:
\begin{verbatim}
<person name="john" surname="smith">
  <info>
   Some text
   <note/>
  </info>
</person>
<inform>
  Another text
  <note/>
</inform>
<person>
    <information>
      Some text
      <note/>
    </information>
</person>
\end{verbatim}
We examine info-like named elements.
If we cluster elements sipmly by their name, we get one cluster with info element, another cluster with inform element and another one with information element.
As \cite{1802522} suggests, we should consider element content in clustering process.
If we take care of element content, elements info, inform and information would look same (in means of node tree inside them - text node and element note).
Elements information and info have even same context (inside person element).

@@

We provide \code{Clusterer} interface for classes to implement.
Its purpose is to cluster bunch of elements (rules) on input, into  bunch of cluster class instances (cluters) on output.
It has methods \code{add()} and \code{addAll()} for adding items for clustering.
Center is method \code{cluster()}, which does the clustering itself.
As it may be time-consuming operation, method throws \code{InterruptedException}.
Implementation should take care of checking whether thread is user interrupted (see \cite[p. 12]{arch}).
After clustering, implementation should hold clusters in member, as it it will be further asked by calling method \code{T getRepresentantForItem(T item)}.
Given item to this method, one can ask for representant of cluster, to which the item belongs.
If no such cluster exists (item was not added for clustering before), we recommend you throwing an exception rather than returning \code{null}.
Missing item will probably  indicate error in algorithm rather than normal workflow.
One can pull clusters from clusterer by calling \code{getClusters()} method.

@@

\nocite{*}
\bibliographystyle{alpha}
\bibliography{literature}

>>>>

\newcommand{\myscale}{0.74}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\jmodule}[1]{\texttt{\textit{#1}}}

@@

\noindent \begin{tabular}{|l|l|} \hline
Responsible developer: & anti \\ \hline
Required tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.RuleDisplayer \\ \hline
Provided tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.inference.Simplifier \\ \hline
Module dependencies:   & AutoEditor, Base, JUNG, Lookup \\ \hline
Public packages:       & \\ \hline
\end{tabular}

\section{Using of factory pattern}
\begin{figure}
  \centering
	\subfigure[Same instance of class returned by NetBeans in two successive inference runs]
	{
	  \centering\includegraphics[scale=\myscale]{factory_pattern1}
    \label{factory_pattern1}
  }
	\subfigure[Factory solution of singleton lookup class problem]
	{
	  \centering\includegraphics[scale=\myscale]{factory_pattern2}
    \label{factory_pattern2}
  }
\end{figure}
\jmodule{TwoStepSimplifier} is divided into three submodules, which are then divided into submodules and so on.
Each submodule is simply class, properly anotated.
User selects proper classes - submodules to work in chain on inference.
Classes are then, in runtime looked up by using NetBeans lookup mechanism (see \cite[p. 16]{archdoc}).
But NetBeans holds one instance of each class, that may be looked upon.
When user runs inference first time, classes are looked up and used.
Maybe class members are initialised, instances are in some state after inference completes.
Then, user clicks run button again, and same instances of classes are returned by lookups.
Oops, theese are not freshly created instances, using them as they are may cause harm.
The problem is ilustrated on fig. \ref{factory_pattern1}.

@@

\section{TwoStep module}
\jmodule{TwoStepSimplifier} is inspired by \cite{1802522} design. Inference proceeds in two steps:
\begin{enumerate}
	\item clustering of element instances into clusters of (probably) same elements
	\item infering regular expression for each element from examples of element contents taken from all elements in cluster
\end{enumerate}
Task of clustering is dedicated to \jmodule{Clusterer} submodule, and task of infering regular expression for each cluster is dedicated to \jmodule{ClusterProcessor} submodule. We will examine both of them.
There is also third submodule called \jmodule{RegularExpressionCleaner} actually, its purpose is just to beatify output regular expressions, no inference logic is implemented there. 
Modules are drawn on fig. \ref{twostep_modules}.
\begin{figure}
	\centering\includegraphics[scale=\myscale]{twostep_modules}
	\caption{Submodules of TwoStep simplifier} \label{twostep_modules}
\end{figure}
We provide one \jmodule{Clusterer} and 4 \jmodule{ClusterProcessor} implementations.
Each of those will be explained further in this document.

\jmodule{TwoStepSimplfier} is implemented in package \code{cz.cuni.mff.ksi.jinfer.twostep}.
Module is implemented in two classes: \code{TwoStepSimplifier} (main logic), \code{TwoStepSimplifierFactory} (lookups, interface to other modules).
Factory class implements \code{Simplifier} interface (defined in package \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference}).
Thus has main method called \code{start} which receives Initial Grammar in form of
\begin{verbatim}
final List<Element> grammar
\end{verbatim}
Each grammar rule is represented as class \code{Element}, where element is left side of rule and it's \code{getSubnodes()} method
returns regular expression, which is right side of rule.
Second parameter is 
\begin{verbatim}
final SimplifierCallback callback
\end{verbatim}
Callback which should be called when work is done.
On output, simplifier provides simplified grammar as a parameter into callback function:
\begin{verbatim}
void finished(final List<Element> grammar);
\end{verbatim}
Thus simplifier receives initial grammar and returns simplified grammar, both represented same way - as elements.

\code{TwoStepSimplifierFactory} in method \code{start()} created new \code{TwoStepSimplifier} class instance, providing three factories of submodules to its constructor.
Then it calls \code{simplify(initialGrammar)} method on this instance and its result passes as parameter to callback function.
Thats all, the magic comes in \code{TwoStepSimplifier} class.

Method \code{TwoStepSimplifier.simplify()} does basically this:
\begin{verbatim}
// 1. cluster elements
final Clusterer<AbstractStructuralNode> clusterer= clustererFactory.create();
clusterer.addAll(initialGrammar);
clusterer.cluster();

@@

// 3. process rules
final ClusterProcessor<AbstractStructuralNode> processor =
             clusterProcessorFactory.create();

for (Cluster<AbstractStructuralNode> cluster : clusterer.getClusters()) {
  final AbstractStructuralNode node =  
                          processor.processCluster(clusterer, cluster.getMembers());
  final RegularExpressionCleaner<AbstractStructuralNode> cleaner = 
                          regularExpressionCleanerFactory.<AbstractStructuralNode>create();
  // 4. add to rules
  finalGrammar.add(
              new Element(node.getContext(),
              node.getName(),
              node.getMetadata(),
              cleaner.cleanRegularExpression(((Element) node).getSubnodes()),
              attList));
}

return finalGrammar;
\end{verbatim}
It creates clusterer, gives it all rules, orders it to cluster elements.
Then, empty list of elements is created as final (simplified) grammar.
For each input rule in initialGrammar, submodule \jmodule{ClusterProcessor} is called to do infering of regular expression of that element.
Finally, regular expression cleaning is done in submodule and new \code{Element} instance is created as a copy of processed node, but with cleaned regexp.

Now we will examine submodules for clustering, processing and cleaning.

TODO sentinel cant be on left side
TODO sentinel processing

\subsection{Clusterer module}
Lets start by example, let input document(s) contain XML:
\begin{verbatim}
<person name="john" surname="smith">
  <info>
    Some text
    <note/>
  </info>
</person>
<inform>
  Another text
  <note/>
</inform>
<person>
    <information>
      Some text
      <note/>
    </information>
</person>
\end{verbatim}
We examine info-like named elements.
If we cluster elements sipmly by their name, we get one cluster with info element, another cluster with inform element and another one with information element.
As \cite{1802522} suggests, we should consider element content in clustering process.
If we take care of element content, elements info, inform and information would look same (in means of node tree inside them - text node and element note).
Elements information and info have even same context (inside person element).

@@

We provide \code{Clusterer} interface for classes to implement.
Its purpose is to cluster bunch of elements (rules) on input, into  bunch of cluster class instances (cluters) on output.
It has methods \code{add()} and \code{addAll()} for adding items for clustering.
Center is method \code{cluster()}, which does the clustering itself.
As it may be time-consuming operation, method throws \code{InterruptedException}.
Implementation should take care of checking whether thread is user interrupted (see \cite[p. 12]{archdoc}).
After clustering, implementation should hold clusters in member, as it it will be further asked by calling method \code{T getRepresentantForItem(T item)}.
Given item to this method, one can ask for representant of cluster, to which the item belongs.
If no such cluster exists (item was not added for clustering before), we recommend you throwing an exception rather than returning \code{null}.
Missing item will probably  indicate error in algorithm rather than normal workflow.
One can pull clusters from clusterer by calling \code{getClusters()} method.

@@

%\nocite{*}
\newpage
\bibliographystyle{alpha}
\bibliography{literature}
